---
title: "Beyond stm: faSTM's extensions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Beyond stm: faSTM's extensions}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.2,
                      dpi = 96, message = FALSE, warning = FALSE)
set.seed(2138)
```

The [companion vignette](faSTM.html) shows that faSTM runs the *same* analysis as
the `stm` package. This one covers what faSTM adds **on top of** that framework:
a richer, more honest `estimateEffect()`, multiple content covariates, alternative
coherence metrics, and a tidyverse-friendly surface. None of these require
leaving the `stm`-compatible object — they are extra tools, not a different model.

The slower fits below are shown with `eval = FALSE` (the code is exactly what you
would run); the lightweight illustrations run live on one base model.

```{r}
library(faSTM)
data(poliblog)
out <- list(documents = poliblog$documents, vocab = poliblog$vocab, meta = poliblog$meta)
```

We fit one prevalence model and reuse it throughout.

```{r basefit}
fit <- stm(out$documents, out$vocab, K = 20,
           prevalence = ~ rating + s(day), data = out$meta,
           verbose = FALSE)
```

## Multiple content covariates

`stm` allows a single content (SAGE) covariate. faSTM accepts several and crosses
them into a **saturated** content model — one topic-word distribution per
combination of levels. The bundled poliblog has two categorical fields, `rating`
(2 levels) and `blog` (6 levels), so `~ rating + blog` yields a 2 × 6 = 12-group
content model:

```{r content, eval=FALSE}
fitC <- stm(out$documents, out$vocab, K = 20,
            prevalence = ~ rating + s(day),
            content    = ~ rating + blog,     # crossed -> 12 SAGE groups
            data = out$meta)
#> faSTM: crossing 2 content covariates (rating, blog) into a saturated
#>        content model with 12 groups.

# stm's SAGE machinery works unchanged on the crossed model:
sageLabels(fitC, n = 5)
labelTopics(fitC, topics = 3)
```

The crossed levels are recorded on the fit (`settings$covariates$contentvars`,
`contenttable`) so per-covariate marginals can be recovered later.

## Honest `estimateEffect()`: weights and cluster-robust SEs

faSTM's `estimateEffect()` keeps `stm`'s method-of-composition uncertainty
(re-drawing topic proportions per simulation) and adds two design features that
real survey/grouped data usually need.

**Survey weights** enter the per-draw regression as weighted least squares.
(poliblog has no design weights; we use an illustrative vector.)

```{r weights}
w <- with(out$meta, ifelse(rating == "Liberal", 1.3, 0.8))   # illustrative
eff_w <- estimateEffect(1:20 ~ rating + s(day), fit,
                        metadata = out$meta, weights = w, nsims = 25)
```

**Cluster-robust standard errors** (here clustering by `blog`) replace the
classical vcov with a sandwich estimator carrying `stm`'s finite-sample
correction:

```{r cluster}
eff_cl <- estimateEffect(1:20 ~ rating + s(day), fit,
                         metadata = out$meta, cluster = out$meta$blog, nsims = 25)
summary(eff_cl, topics = 3)
```

`summary()` also takes a `p.adjust.method` (e.g. `"BH"`) to correct across topics,
and reports per-equation R² and F diagnostics.

## Random effects in prevalence

Prevalence formulas may include `lme4`-style random-effect terms; faSTM fits a
mixed model per posterior draw and pools with Rubin's rules. Useful when
documents nest in groups (authors, outlets, years):

```{r ranef, eval=FALSE}
eff_re <- estimateEffect(1:20 ~ rating + (1 | blog), fit,
                         metadata = out$meta, nsims = 25)
summary(eff_re, topics = 3)   # fixed effects + pooled variance components
```

## Average marginal effects

Coefficients on splines or interactions are hard to read. `ame()` turns an
estimated effect into an interpretable average marginal effect — the mean change
in topic proportion for a one-unit (or factor-level) change in a covariate,
averaged over the sample:

```{r ame}
eff <- estimateEffect(1:20 ~ rating + s(day), fit, metadata = out$meta, nsims = 25)
ame(eff, covariate = "rating", topics = c(3, 7, 20))
```

## Coherence: NPMI and C_V

Alongside Mimno's semantic coherence, faSTM offers the two coherence metrics most
used in the modern topic-model literature — **NPMI** and **C_V**:

```{r coh}
data.frame(
  topic = 1:5,
  mimno = round(coherence(fit, measure = "mimno", M = 10)[1:5], 2),
  npmi  = round(coherence(fit, measure = "npmi",  M = 10)[1:5], 3),
  c_v   = round(coherence(fit, measure = "c_v",   M = 10)[1:5], 3)
)
```

`search_k()` can select K by any of these (and parallelizes across K):

```{r searchk, eval=FALSE}
co <- as_corpus(poliblog)
search_k(co, K = c(10, 20, 30), prevalence = ~ rating + s(day),
         measure = "npmi", cores = 3)
```

## A tidyverse-friendly surface

faSTM ships `broom` generics, so model output flows straight into `dplyr`/`ggplot2`.

```{r broom}
# tidy the topic-word matrix (also matrix = "frex" or "gamma")
head(tidy(fit, matrix = "beta"), 4)

# one-row model summary
glance(fit)

# tidy an estimated effect into a coefficient table
head(tidy(eff), 4)
```

`predict()` infers topic proportions for new documents against the fitted model:

```{r predict, eval=FALSE}
predict(fit, newdata = held_out_corpus)   # a faSTM_corpus / dfm of new docs
```

## Summary

Everything here returns ordinary data frames or `stm`-compatible objects, so it
composes with the rest of your analysis. The companion vignette covers the
`stm`-equivalent workflow; this one is the extra mile faSTM adds on top.
