---
title: "faSTM: the stm vignette, run on faSTM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{faSTM: the stm vignette, run on faSTM}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.2,
                      dpi = 96, message = FALSE, warning = FALSE)
set.seed(2138)
```

This vignette walks through the **same analysis as the `stm` package vignette**
(Roberts, Stewart & Tingley), using the identical CMU 2008 political-blog corpus
— but every model is **fit live** here, because faSTM fits in seconds where `stm`
takes minutes. (The `stm` vignette loads pre-computed objects to avoid the wait;
faSTM does not need to.) The code mirrors the `stm` vignette's calls. Because
each fit is fresh, the *topic numbers* differ from the original — the workflow,
not the specific topics, is what carries over.

```{r}
library(faSTM)
```

## Ingesting data

faSTM reads prepared text from `quanteda`/`tidytext` rather than tokenizing
itself. A typical preparation:

```{r prep, eval=FALSE}
library(quanteda)
dfmat <- corpus(my_data, text_field = "documents") |>
  tokens(remove_punct = TRUE) |>
  tokens_remove(stopwords("en")) |>
  dfm() |>
  dfm_trim(min_termfreq = 5)
corpus <- as_corpus(dfmat)        # quanteda docvars become the metadata
```

For this vignette we use the bundled poliblog corpus (the `stm` vignette's
`poliblog5k`), already prepared:

```{r data}
data(poliblog)
poliblog
out <- list(documents = poliblog$documents, vocab = poliblog$vocab, meta = poliblog$meta)
```

## Estimating the structural topic model

The headline call mirrors the `stm` vignette exactly — topic prevalence varies
with `rating` and a smooth function of `day`:

```{r fit}
poliblogPrevFit <- stm(out$documents, out$vocab, K = 20,
                       prevalence = ~ rating + s(day), data = out$meta,
                       init.type = "Spectral", seed = 2138)
```

That fit took seconds, not minutes.

## Model selection and search

`selectModel()` fits several models from different initializations and keeps the
ones on the semantic-coherence / exclusivity frontier; `plotModels()` shows them.
(Reduced to a few candidates here to keep the vignette quick.)

```{r select, fig.height=4}
poliblogSelect <- selectModel(out$documents, out$vocab, K = 20, N = 5,
                              prevalence = ~ rating + s(day), data = out$meta, seed = 2138)
plotModels(poliblogSelect)
```

`searchK()` sweeps the number of topics, reporting held-out likelihood, semantic
coherence and exclusivity — and parallelizes across K:

```{r searchk, fig.height=4.2}
storage <- searchK(out$documents, out$vocab, K = c(10, 20),
                   prevalence = ~ rating + s(day), data = out$meta, cores = 2)
plot(storage)
```

## Interpreting topics

Top words by probability, FREX, lift and score:

```{r labels}
labelTopics(poliblogPrevFit, c(3, 7, 20))
```

Representative documents per topic, displayed as wrapped quotes:

```{r thoughts, fig.height=3.2}
thoughts3 <- findThoughts(poliblogPrevFit, texts = out$meta$text, n = 2, topics = 3)$docs[[1]]
plotQuote(substr(thoughts3, 1, 200), width = 50, main = "Topic 3")
```

Topics ranked by their expected prevalence in the corpus:

```{r summary}
plot(poliblogPrevFit, type = "summary")
```

## Covariate effects on topic prevalence

`estimateEffect()` regresses topic proportions on the covariates, propagating
topic-estimation uncertainty (the honest method of composition):

```{r effect}
out$meta$rating <- as.factor(out$meta$rating)
prep <- estimateEffect(1:20 ~ rating + s(day), poliblogPrevFit,
                       meta = out$meta, uncertainty = "Global")
summary(prep, topics = 1)$tables[[1]]
```

Difference in topic prevalence between Liberal and Conservative blogs:

```{r diff, fig.height=3.6}
plot(prep, covariate = "rating", topics = c(3, 7, 20), model = poliblogPrevFit,
     method = "difference", cov.value1 = "Liberal", cov.value2 = "Conservative",
     xlab = "More Conservative ... More Liberal")
```

A topic's prevalence over time (smooth term in `day`):

```{r cont, fig.height=3.6}
plot(prep, "day", method = "continuous", topics = 7, model = poliblogPrevFit)
```

## Topical content

Letting word *use within topics* vary by `rating` (a SAGE content covariate),
then comparing the two sides' vocabulary for a topic:

```{r content, fig.height=5}
poliblogContent <- stm(out$documents, out$vocab, K = 20,
                       prevalence = ~ rating + s(day), content = ~ rating,
                       data = out$meta, init.type = "Spectral", seed = 2138)
plot(poliblogContent, type = "perspectives", topics = 1)
```

Comparing the vocabulary of two topics:

```{r persp2, fig.height=5}
plot(poliblogPrevFit, type = "perspectives", topics = c(12, 20))
```

## Interactions

Prevalence can interact covariates — here `rating` with time — and the effect
plot can condition on a moderator value:

```{r interaction, fig.height=3.8}
poliblogInteraction <- stm(out$documents, out$vocab, K = 20,
                           prevalence = ~ rating * day, data = out$meta,
                           init.type = "Spectral", seed = 2138)
prepInt <- estimateEffect(c(16) ~ rating * day, poliblogInteraction,
                          metadata = out$meta, uncertainty = "None")
plot(prepInt, covariate = "day", model = poliblogInteraction, method = "continuous",
     xlab = "Days", moderator = "rating", moderator.value = "Liberal", topics = 16)
```

## More visualization

A word cloud for a topic, the topic-correlation network, and the convergence
trajectory:

```{r viz, fig.height=4.5}
cloud(poliblogPrevFit, topic = 7)
plot(topicCorr(poliblogPrevFit))
plot(poliblogPrevFit$convergence$bound, type = "l",
     ylab = "Approximate Objective", main = "Convergence")
```

## Out-of-sample documents

New documents get topic proportions by holding the fitted topics fixed:

```{r newdocs}
theta_new <- fit_new_documents(poliblogPrevFit, poliblog)
dim(theta_new)
```

---

Everything above is the `stm` vignette's workflow, run on faSTM: the same
function names and arguments, the same corpus, the same kinds of plots — fit in
seconds and with an honest `estimateEffect`. Existing `stm` scripts port with
little more than the changes shown here.
