---
title: "Validation: parity with stm, and fit quality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation: parity with stm, and fit quality}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Runs only when the stm package is available (it is a Suggests dependency); the
# whole point of this article is to compare against live stm output.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      message = FALSE, warning = FALSE,
                      eval = requireNamespace("stm", quietly = TRUE))
set.seed(2138)
```

faSTM is a reimplementation of the Structural Topic Model, not a wrapper around
`stm`. It has its own optimizer (a Rust variational EM), so two questions decide
whether you can trust it:

1. **Given the same fitted model, do faSTM's post-fit numbers match `stm`'s?**
2. **Do faSTM's own fits reach the same quality as `stm`'s**, even though the
   topic decomposition differs?

This article answers both by running `stm` live and comparing. Every check is
guarded with `stopifnot()`, so this page fails to build if parity ever breaks.

```{r data}
library(faSTM)
library(stm)
data(poliblog)
out <- list(documents = poliblog$documents, vocab = poliblog$vocab, meta = poliblog$meta)
```

## Same model, same numbers

The fitted object is `stm`-shaped, so `stm`'s own readers run on a faSTM fit. We
fit once with faSTM, then compute every inspection metric **both ways** (with
faSTM's functions and with `stm`'s) on that single shared model. Identical
inputs should give identical outputs.

```{r fit}
fit <- faSTM::stm(out$documents, out$vocab, K = 20,
                  prevalence = ~ rating + s(day), data = out$meta,
                  init.type = "Spectral", seed = 2138, verbose = FALSE)
```

### Topic labels (probability, FREX, lift, score)

```{r labels}
fa <- faSTM::label_topics(fit, n = 7)
sl <- stm::labelTopics(fit, n = 7)
same <- vapply(c("prob", "frex", "lift", "score"),
               function(m) identical(unname(fa[[m]]), unname(sl[[m]])), logical(1))
same
stopifnot(all(same))
```

All four rankings return the identical words, topic by topic.

### Semantic coherence and exclusivity

```{r quality}
coh_diff  <- max(abs(faSTM::semantic_coherence(fit, M = 10) -
                     stm::semanticCoherence(fit, documents = out$documents, M = 10)))
excl_diff <- max(abs(faSTM::exclusivity(fit, M = 10) -
                     stm::exclusivity(fit, M = 10)))
c(max_coherence_diff = coh_diff, max_exclusivity_diff = excl_diff)
stopifnot(coh_diff < 1e-8, excl_diff < 1e-8)
```

Both agree to floating-point precision: faSTM's `inspect.R` ports `stm`'s
`semCoh1beta` / `js.estimate` / FREX formulas directly.

## Different fit, comparable quality

faSTM's optimizer is not `stm`'s, and the STM objective is non-convex, so an
independent run settles into its own optimum with its own topic numbering. The
question is whether that optimum is as *good*. The fair, engine-neutral test is
held-out predictive likelihood by document completion: hold out half the tokens
in each document, fit on the rest, and score the held-out tokens. We run the same
held-out set through both packages.

```{r heldout}
ho <- stm::make.heldout(out$documents, out$vocab, seed = 2138)

ff <- faSTM::stm(ho$documents, out$vocab, K = 20, prevalence = ~ rating + s(day),
                 data = out$meta, init.type = "Spectral", seed = 2138, verbose = FALSE)
sf <- stm::stm(ho$documents, out$vocab, K = 20, prevalence = ~ rating + s(day),
               data = out$meta, init.type = "Spectral", seed = 2138, verbose = FALSE)

ll_faSTM <- mean(stm::eval.heldout(ff, ho$missing)$expected.heldout)
ll_stm   <- mean(stm::eval.heldout(sf, ho$missing)$expected.heldout)

data.frame(
  engine     = c("faSTM", "stm"),
  heldout_LL = round(c(ll_faSTM, ll_stm), 4),
  iterations = c(ff$convergence$its, length(sf$convergence$bound)))
```

```{r heldout-check}
rel_gap <- abs(ll_faSTM - ll_stm) / abs(ll_stm)
round(100 * rel_gap, 3)               # percent difference in held-out likelihood
stopifnot(rel_gap < 0.02)             # within 2%
```

The two fits land within a fraction of a percent on held-out likelihood, so the
optima are of comparable quality. faSTM reaches it in more iterations, but each
iteration is cheaper, so it still converges faster in wall-clock time (see the
[Get started](faSTM.html) article).

## What this means for your analysis

- **Post-fit numbers are safe to compare.** Labels, coherence, and exclusivity
  computed by faSTM equal `stm`'s for any given model.
- **Fits are not identical, by design.** faSTM and `stm` find different (equally
  good) topic decompositions. For a result that must survive replication, fit your
  final, reported model in whichever package your readers will rerun, and report
  the package and version. This page is the evidence that either choice is sound.
